---
title: "Assignment 1"
author: "Eswar Dumpa"
date: "2024-02-04"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Importing Dataset
```{r import}
diabetes <- read.csv("diabetes.csv")
head(diabetes)
```

#Importing data from kaggle dataset using the below link
https://www.kaggle.com/datasets/akshaydattatraykhare/diabetes-dataset

## Descriptive Statistics

```{r descriptive}
summary(diabetes)
```
#it displays the summary statistics of the "diabetes" Dataset.The Dataset contains 768 observations.

## Transform

```{r transform}

library(dplyr)
diabetes_2<- mutate(diabetes,Outcome=case_when(
  Outcome==1 ~ "Positive",
  Outcome==0 ~ "Negative"
))
diabetes_2$Outcome <- as.factor(diabetes_2$Outcome)
summary(diabetes_2)
```
#here the numeric variable "outcome" is transposed into categorical variable, any value greater than zero is assigned as positive and else negative.


## Plot

```{r plot}
plot(diabetes_2$Glucose)
```

#it creates a scatter plot of variable "glucose" levels of the patients, the max glucose level is 199.0 in the data set.



